---
title: "Exam 2"
execute:
  echo: true
  warning: false
  error: false

format: 
  html: 
    code-fold: true
---

# Background

```{r}
library(tidytuesdayR)
library(tidyverse)
library(rnaturalearth)
library(sf)
library(dplyr)

tuesdata <- tt_load('2020-02-18')
fc <- tuesdata$food_consumption

str(fc)
head(fc, 22)
```
```{r}
# Inspect country variable: The dataframe contains information about 130 countries

fc |>
  count(country)

# Inspect food_category variable 
fc |>
  select(food_category)

# Fix food_category variable 

fc <- fc |>
  mutate(food_category = fct_recode(food_category,
                                    "Lamb" = "Lamb & Goat",
                                    "Dairy" = "Milk - inc. cheese",
                                    "Wheat" = "Wheat and Wheat Products",
                                    "Nuts" = "Nuts inc. Peanut Butter"))

```
# Most Consuming Countries 

::: {.callout-note title="Research Question 1"}

Which 5 countries consume the most food are?

 - Show the answer as a 2-column table and as graph.
 - Sort the countries based on consumption from largest to smallest 
 - Use the cleaned dataframe
 
:::

```{r}
top5_consumers <- fc |>
  group_by(country) |>
  summarise(total_consumption = sum(consumption, na.rm = TRUE)) |>
  arrange(desc(total_consumption)) |>
  head(5)

top5_consumers
```
```{r}
ggplot(top5_consumers, aes(x = reorder(country, total_consumption), y = total_consumption, fill = country)) +
  geom_col(show.legend = FALSE) +
  labs(title = "Top 5 Countries by Total Food Consumption",
       x = "Country",
       y = "Total Consumption (kg/person/year)") +
  theme_minimal()
```
# Most Consuming Countries of Each Food

::: {.callout-note title="Research Question 2"}

Which top 5 countries consume each food are?

  - Show the answer as a 3-column table (show the food category as the first column followed by country and consumption) and as a graph.
  - For each food, sort the countries based on consumption from largest to smallest.
  - Use the cleaned dataframe
 
:::

```{r}
top5_per_food <- fc |>
  group_by(food_category, country) |>
  summarise(total_consumption = sum(consumption, na.rm = TRUE)) |>
  arrange(food_category, desc(total_consumption)) |>
  mutate(rank = row_number()) |>
  filter(rank <= 5) |>
  select(food_category, country, total_consumption)

top5_per_food
```
```{r}
#| fig-height: 10
#| fig-width: 12
ggplot(top5_per_food, aes(y = reorder(country, total_consumption),
                          x = total_consumption, fill = country)) +
  geom_col(show.legend = FALSE) +
  facet_wrap(~ food_category, scales = "free_y") +
  labs(title = "Top 5 Countries by Food Category Consumption",
       x = "Country",
       y = "Consumption (kg/person/year)") +
  theme_minimal()
```

# Food Consumption 

::: {.callout-note title="Research Question 3"}

What does the consumption of each food look like?

 - show a choropleth map for each food.
 - Use the cleaned dataframe. 
 
:::

## Viz: Attempt 1

```{r}
ne_countries(returnclass = "sf") |>
  select(name, geometry) |>
  left_join(fc |> select(-co2_emmission),
            join_by(name == country)) |>
  ggplot() +
  geom_sf(aes(fill = consumption)) +
  facet_wrap(~food_category) +
  theme(legend.position = "bottom")
```

What problems does the above visualization suffer?
 - Unclear or meaningless NA facet. 
 - The shading is very dark for most countries across facets, making hard to distinguish differences in consumption
 - There are no country borders, making harder to distinguish regions
 - Some countries are missing, posible because when joining data sets, the names were different spelled. 
 
## Viz: Attempt 2 

```{r}
ne_countries(returnclass = "sf") |>
  select(name, geometry) |>
  mutate(name = ifelse(name == "United States of America", "USA", name)) |>
  mutate(name = ifelse(name == "Bosnia and Herz.", "Bosnia and Herzegovina", name)) |>
  mutate(name = ifelse(name == "Czechia", "Czech Republic", name)) |>
  mutate(name = ifelse(name == "Taiwan", "Taiwan. ROC", name)) |>
  left_join(fc |> select(-co2_emmission),
            join_by(name == country)) |>
  pivot_wider(names_from = food_category,
              values_from = consumption) |>
  select(-"NA") |>
  pivot_longer(cols = c(-name, -geometry),
               names_to = "food_category",
               values_to = "consumption") |>
  ggplot() +
  geom_sf(aes(fill = consumption)) +
  facet_wrap(~food_category) +
  theme(legend.position = "bottom")
```

What problems does the above visualization suffer?
 - The color palette doesn't differentiate well between low and high consumption. It is also no inclusive for colorblind viewers.
 - There are no countries labels or border lines. 
 - Many small maps are hard to read and compared. 
 
## Viz: Attempt 3

```{r}
ne_countries(returnclass = "sf") |>
  select(name, geometry) |>
  mutate(name = ifelse(name == "United States of America", "USA", name)) |>
  mutate(name = ifelse(name == "Bosnia and Herz.", "Bosnia and Herzegovina", name)) |>
  mutate(name = ifelse(name == "Czechia", "Czech Republic", name)) |>
  mutate(name = ifelse(name == "Taiwan", "Taiwan. ROC", name)) |>
  left_join(
    fc |> 
      select(-co2_emmission) |>
      group_by(food_category) |>
      mutate(consumption = (consumption - mean(consumption))/sd(consumption)),
      join_by(name == country)) |>
  pivot_wider(names_from = food_category, values_from = consumption) |>
  select(-"NA") |>
  pivot_longer(cols = c(-name, -geometry),
               names_to = "food_category",
               values_to = "consumption") |>
  ggplot() +
  geom_sf(aes(fill = consumption)) +
  facet_wrap(~food_category) +
  theme(legend.position = "bottom")
```
How can the visualization be enhanced?
 - Color scale improvements 
 - Small countries can be hard to see on a world map
 - Countries outlines/borders could help visualization 
