---
title: "Solo Project"
execute:
  echo: true
  warning: false
  error: false

format: 
  html: 
    code-fold: true
---

```{r}
library(ggplot2)
library(dplyr)
library(readr)
library(tidyverse) 
library(maps)
library(ggthemes)

elections_project <- read_csv("~/Documents/GitHub/portfolio-kherrera23/data/Election.csv")

# Filter for Minnesota
mn_election <- elections_project |>
  filter(state == "Minnesota") |>
  select(county, trump16, clinton16, otherpres16)

# Determine the winning candidate per county
mn_winners <- mn_election |>
  mutate(winner = case_when(
    trump16 > clinton16 & trump16 > otherpres16 ~ "Trump",
  clinton16 > trump16 & clinton16 > otherpres16 ~ "Clinton", 
  TRUE ~ "Other"
  ))
```
```{r}
#| fig-height: 8
#| fig-width: 12
# Minnesota county map
library(tigris)
options(tigris_use_cache = TRUE)
mn_map <- counties(state = "MN", class = "sf")
mn_map <- mn_map |>
  mutate(county = str_remove(NAME, "County"))

# Merge election data with county map
mn_map <- mn_map |>
  left_join(mn_winners, by = "county")

ggplot(mn_map) +
  geom_sf(aes(fill = winner), color = "white") +
  scale_fill_manual(values = c("Trump" = "red", "Clinton" = "blue", "Other" = "gray")) +
  labs(title = "Minnesota 2016 Presidential Election Results (by county)",
       subtitle = "Candidate with the most votes in each county - not proportional to population",
       fill = "Winning Candidate",
       caption = "Source: MIT Election Data and Science Lab, Harvard Dataverse (2019)
       Author: Kirsty Herrera Garcia") +
  theme_map() +
    theme_map() +
  theme(
    plot.title = element_text(size = 20, face = "bold"),
    plot.subtitle = element_text(size = 14),
    legend.title = element_text(size = 14),
    legend.text = element_text(size = 12),
    plot.caption = element_text(size = 10)
  )
```

::: {.callout-note title="Summary"}

This visualization shows the results of the 2016 presidential election per county in the state of Minnesota. This data was accessed via Harvard Dataverse, a publicly available database housing this 2019 study by the Massachusetts Institute of Technology in Cambridge, Massachusetts, where MIT Election Data and Science Lab collects, examines, and distributes election data. At the same time, the data was curated from the American Community Survey and election results. It is a collection of demographic, past federal, and past state election data for counties in the United States from 2012 to 2016. The missing values from two counties were not registered in the data set but there is not explanation as why that happened. 

This maps shows the "winning" candidate per county, meaning the candidate with majority of total votes. Although it can be observed a majority red, or counties were Trump was more voted than Clinton, Minnesota was a "blue state" in 2016, meaning that the 10 electoral votes of the state were given to Hillary Clinton. Without knowing the electoral system of the US, this visualization could be misleading regard the political support of the state in the 2016 presidential election. It is important to note that more populated counties are blue, which also seems to relate to their geographical location in main cities of the state. 

:::

